\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}   % Images
\usepackage{subfiles}   % Include subfiles
\usepackage{hyperref}   % Nice referencing with clickable links
\usepackage{amsmath}    % Math
%\usepackage{biblatex}   % Citations
\usepackage[english]{babel}
\usepackage{csquotes}   % Inline and display quotations
\usepackage[nameinlink,noabbrev]{cleveref}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[margin=1.5in]{geometry}
\graphicspath{{figures/}{../figures/}}
\usepackage[parfill]{parskip}
\usepackage{float}
\usepackage{breakurl}
\usepackage{siunitx}
\usepackage{xcolor}

\title{}
\author{Daniel Pinjusic, Gudmund Gunnarsen and Ida Due-Sørensen}
\date{\today}

\begin{document}

\maketitle

\subfile{sections/abstract}

\subfile{sections/introduction}

\subfile{sections/theory}

\subfile{sections/method}

\subfile{sections/results}

\subfile{sections/discussion}

\subfile{sections/conclusion}

\section{Appendix}

Github repository with codes and figures can be found at \burl{https://github.com/idadue/ComputationalPhysics/tree/master/project3}.


\begin{thebibliography}{9}
\bibitem{Taut}
Taut, M. \textit{Two electrons in an external oscillator potential: Particular analytic solutions of a Coulomb potential correlation problem} Physical Review A, Vol. 48, number 5, 1993.

\bibitem{matrix-computations}
Van Loan, C. F. and Golub, G. H.: \textit{Matrix Computations}, chapter 8.5. The John Hopkins University Press, Baltimore, Maryland, 2013.

\bibitem{arma1}
C. Sanderson and R. Curtin.
\textit{Armadillo: a template-based C++ library for linear algebra.}
Journal of Open Source Software, Vol. 1, pp. 26, 2016.

\bibitem{arma2}
Sanderson, C. and  Curtin, R.:
\textit{A User-Friendly Hybrid Sparse Matrix Class in C++.}
Lecture Notes in Computer Science (LNCS), Vol. 10931, pp. 422-430, 2018. 

\bibitem{project2}
Hjort-Jensen, M: Project2,
\\\texttt{\burl{https://github.com/CompPhysics/ComputationalPhysics/blob/master/doc/Projects/2020/Project2/pdf/Project2.pdf}}

\bibitem{compphy} 
Hjort-Jensen, M: Computational Physics,
%\\\texttt{\href{https://github.com/CompPhysics/ComputationalPhysics/blob/master/doc/Lectures/lectures2015.pdf}{Find on course Github: ./doc/lectures/}}
\\\texttt{\burl{https://github.com/CompPhysics/ComputationalPhysics/blob/master/doc/Lectures/lectures2015.pdf}}




\end{thebibliography}
% \printbibliography

\end{document}
