\documentclass[../main.tex]{subfiles}

\begin{document}
\section{Introduction}\label{sec:introduction}

%The purpose of this project is to simulate the solar system and the path of celestial objects. 

The solar system is a complicated system with many interacting bodies, and it is impossible to solve analytically for instance the path of every object in the solar system. However, this complexity allows for a numerical approach in the solutions. In particular, the planetary motions in the solar system can be modelled as coupled differential equations, and these equations can be written in general forms, such that the solutions can be found using general numerical methods. Differential equations are ubiquitous in physics, appearing in almost every field in some fashion, and so the methods used in this project have wide-ranging applications.

Another important aspect of this project is object oriented programming. We will work to generalize the code used so that it may be reused in different parts of both this project and possibly another. The system looked at here is particularly suited to object oriented programming as we need to work on multiple sets of systems, with just Earth-Sun, Earth-Sun-Jupiter, the entire Solar system, the Mercury-Sun system with relativistic corrections, and the Earth-Sun system with a generalized force of gravity. Rewriting code for each and every one of these systems would take a large amount of time, and also be prone to errors. Furthermore, if the code is generalized enough, we could reuse large parts of it for other similar systems, such as molecular dynamics where the forces are also conservative and therefore of a similar form.

We will solve first the Earth-Sun system, ignoring all other planets and objects, using the Euler method and the velocity Verlet method, comparing the two solutions and their performance. Then, having made code that can simulate a specific system, testing various quantities such as the escape velocity and conservation of angular momentum to ensure that the algorithms work well, we will extend the framework to a three-body problem, and further to the entire solar system. Finally we check the perihelion precession of Mercury, relating to relativistic calculations of gravity.

%Structure of the report:

First we review some important astronomical and mechanical laws, before explaining the method with which the solar system is simulated. We present images of the simulated solar system, testing escape velocity, different forms of the gravitational force, and the perihelion precession of Mercury with relativistic corrections. We compare the Forward Euler and Velocity Verlet method on stability and energy conservation.

\end{document}
